---
title: "Problem Set 4"
author: "Ashutosh Singh"
date: "Tuesday, August 18, 2015"
output: html_document
---

This file contains the solution for Problem Set 4 of Data Analysis with R course


Load the Libraries and data
```{r}
library(ggplot2)
head(diamonds)
```

 Your first task is to create a
 scatterplot of price vs x.
 using the ggplot syntax.

```{r}
ggplot(data = diamonds, aes(x=x, y = price)) +
  geom_point()
```

What is the corelation between price and x?
```{r}
cor.test(diamonds$price,diamonds$x)
```

What is the corelation between price and y?
```{r}
cor.test(diamonds$price,diamonds$y)
```
What is the corelation between price and z?
```{r}
cor.test(diamonds$price,diamonds$z)
```

Create a simple scatter plot of price vs depth(z)
```{r}
ggplot(data = diamonds, aes(x=z, y =price ))+
  geom_point()+
  coord_cartesian(xlim=c(0,10))  
```

 Change the code to make the transparency of the
 points to be 1/100 of what they are now and mark
 the x-axis every 2 units. See the instructor notes
 for two hints.

```{r}
ggplot(data = diamonds, aes(x=depth, y =price ) )+
  geom_point(alpha=1/100) +
  scale_x_continuous(breaks = seq(40,80,2))  
```

 Create a scatterplot of price vs carat
 and omit the top 1% of price and carat
 values.
 
```{r}
ggplot(data=diamonds,aes(x=carat,y = price)) +
  geom_point() +
  scale_y_continuous(limits=c(quantile(diamonds$price,probs = 0.99), quantile(diamonds$price,probs=1)))
```

 Create a scatterplot of price vs. volume (x * y * z).
 This is a very rough approximation for a diamond's volume.

 Create a new variable for volume in the diamonds data frame.
 This will be useful in a later exercise.
```{r}
diamonds$volume <- diamonds$x*diamonds$y*diamonds$z
ggplot(data = diamonds, aes(x=volume, y = price))+
  geom_point()+
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50))

nrow(subset(diamonds,volume==0))

```


Correlation between price and volume
```{r}
diamonds_subset <- subset(diamonds,diamonds$volume>0 & diamonds$volume<800)
cor.test(diamonds_subset$price,diamonds_subset$volume)

```

 Subset the data to exclude diamonds with a volume
 greater than or equal to 800. Also, exclude diamonds
 with a volume of 0. Adjust the transparency of the
 points and add a linear model to the plot. (See the
 Instructor Notes or look up the documentation of
 geom_smooth() for more details about smoothers.)

```{r}
diamonds_subset <- subset(diamonds,diamonds$volume>0 & diamonds$volume<800)
ggplot(data=diamonds_subset,aes(x=volume,y = price))+
  geom_point(alpha=1/10) +
  ylim(0,20000)+
  stat_smooth(method="lm", color='orange', se=FALSE)

```
 Use the function dplyr package
 to create a new data frame containing
 info on diamonds by clarity.

 Name the data frame diamondsByClarity

 The data frame should contain the following
 variables in this order.

       (1) mean_price
       (2) median_price
       (3) min_price
       (4) max_price
       (5) n

 where n is the number of diamonds in each
 level of clarity.
 
 Load Library dplyr
 
 
```{r}
library(dplyr)
```



```{r}

diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(as.numeric(price)),
            min_price = min(price),
            max_price = max(price),
            n=n()) 
```

We've created summary data frames with the mean price
 by clarity and color. You can run the code in R to
 verify what data is in the variables diamonds_mp_by_clarity
 and diamonds_mp_by_color.
 
 
```{r}

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

```

Your task is to write additional code to create two bar plots
 on one output image using the grid.arrange() function from the package
 gridExtra.

```{r}
library(gridExtra)

p1<-ggplot(data=diamonds_mp_by_clarity, aes(x=clarity, y = mean_price))+
  geom_bar(stat="identity")

p2<-ggplot(data=diamonds_mp_by_color, aes(x=color, y = mean_price))+
  geom_bar(stat="identity")

grid.arrange(p1,p2,ncol=2)
```

The data is counter intuitive as the quality of clarity and color increased the price decreases. LEts look at the mean price  by cut
```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- group_by(diamonds_by_cut,mean_price=mean(price))
ggplot(data = diamonds_mp_by_cut,aes(x=cut,y=mean_price))+
  geom_bar(stat="identity")

```


